Multi-Index and Multi-Level Monte Carlo Evaluation of HJM Models

### 2016-04-25 tests/HJM/manuscript.tex
first runs to the document running, scripts set
generic HJM template
add empirical rates to notes

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{enumerate}
\usepackage{textcomp}
\usepackage{mathrsfs }
\usepackage[section]{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{patchplots}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{multirow}

@@

\begin{abstract}
Abstract here!!!
\end{abstract}

@@

We are interested in analysing the Forward Euler Finite Element evaluation of forward curve models of HJM models. The main approach was represented with a convergence analysis in \cite{bjork2013monte}, whose notation we largely follow.

@@

@@

Due to the fact that the discretisation errors relating to $\tau$
depend on the minimum of $\ell_2$ and $\ell_3$, it seems natural
to assume $\ell_3=\ell_2$. This gives us a two-dimensional
Multi-Index Monte Carlo formulation with the following effective parameters
\begin{align}
\bar \beta &= [2,2]^T ,\\
\bar \gamma &= [1,1]^T , \\
\bar s &= [2^{-1},1], \\
\bar w & = [2,2].
\end{align}

@@

\section{Multi-Index Monte Carlo}

\section{Multi-Level Monte Carlo}

>>>>

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{enumerate}
\usepackage{textcomp}
\usepackage{mathrsfs }
\usepackage[section]{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{patchplots}
%\usepackage{subfigure}
\usepackage{subcaption}

\usepackage{hyperref}
\usepackage{comment}
\usepackage{multirow}

@@

\begin{abstract}
Notes of MIMC and MLMC evaluation of HJM models.
\end{abstract}

@@

\section{Intro}

We are interested in analysing the Forward Euler Finite Difference evaluation of forward curve models of HJM models. The main approach was represented with a convergence analysis in \cite{bjork2013monte}, whose notation we largely follow.

@@

Let our quantity of interest be defined as the expectation of a functional $\mathcal F$
of the forward curve:
\begin{align*}
\mathcal F \parent f =& F \parent{\int_0 ^{t_{max}} f \parent{s,s} ds}
G \parent{ \int_{\tau_a}^{\tau_{max}} \Psi \parent{f \parent{t_{max},\tau}} d\tau }
\\
&+
\int_{0}^{t_{max}} F \parent{\int_0^{s} f \parent{s',s'} ds'} U \parent{f\parent{s,s}} ds.
\end{align*}

@@

Due to the fact that the discretisation errors relating to $\tau$
depend on the minimum of $\ell_2$ and $\ell_3$, it seems natural
to assume $\ell_3=\ell_2$. This gives us a two-dimensional
Multi-Index Monte Carlo formulation with the following effective parameters
\begin{align}
\bar \beta &= [2,2]^T ,\\
\bar \gamma &= [1,1]^T , \\
\bar s &= [1,2], \\
\bar w & = [1,2].
\end{align}

@@

\section{Numerical tests}

For the simplest possible example, let our model be Gaussian with
the following exponential covariance structure
\begin{align}
\expp{df \parent{t,\tau} df \parent{t, \tau'}}
=
\sigma^2 \expf{- \kappa \absval{\tau-\tau'}} dt.
\end{align}

Let our first test case be arguably simplest infinite dimensional one possible,
\begin{align*}
F \parent{x} &= 1-x,\\
U \parent{x} &= 0, \\
G \parent{x} &= x, \\
\Psi \parent{x} &= x, \\
f \parent{0,\tau} &= 0, \\
\kappa &= 2.
\end{align*}

In \ref{fig:empInfDimRates} we get see that we have rate one for both 
the weak and the strong rate along both dimensions for this text example.

\begin{figure}
\includegraphics[scale=0.7]{figure_1.pdf}
\caption{Sample realisation of a forward curve with the discretisation level 
$\overline \ell = [5,5,5]^T$.}
\end{figure}

It remains to deduce the cost of simulating the cost of a realisation. In order to
simulate Gaussian random variables one needs the Cholesky decomposition of the covariance matrix. The cost to generate this $N_\tau \times N_\tau$-matrix is $N_\tau^3$, with $N_\tau = 2^{\ell_3} 2^{\ell_2}$. Using the Cholesky decomposition, the cost to take a a time step becomes $N_\tau^2$. Putting all together, the cost to generate $M$ Monte Carlo realisations with $N_t$ time steps and $N_\tau$ discretisation steps along the maturity direction becomes
\begin{align*}
C \parent{\Delta t, \Delta \tau, M} = N_\tau^3 + M N_\tau^2 N_t.
\end{align*}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{dim_1_strong.pdf}
        %\caption{A gull}
        \label{fig:gull}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{dim_1_weak.pdf}
        %\caption{A tiger}
        \label{fig:tiger}
    \end{subfigure}
    \\
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{dim_2_strong.pdf}
        %\caption{A mouse}
        \label{fig:mouse}
    \end{subfigure}
        \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{dim_2_weak.pdf}
        %\caption{A mouse}
        \label{fig:mouse}
    \end{subfigure}
    \caption{\label{fig:empInfDimRates}Empirical strong and weak convergence rate for the $\Delta t$ and $\Delta \tau$ discretisation errors.}
\end{figure}
